Finite element analysis of multi-dimensional and simplified models for beams

### 2024-04-11 final_seminar/main.tex
speech

\title{Finite element analysis of multi-dimensional and simplified models for beams and plates}
\author{Rudi du Toit}
\date{\today}
\institute{University of Pretoria}

@@

\section{Theory}
    \subsection{General vibration problem}
        \begin{frame}{General vibration problem}
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+a(u'(t),v)+b(u(t),v) = (f(t),v) \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.
        \end{frame}

    \subsection{Existence and uniqueness of solutions}
        \begin{frame}{Existence and uniqueness of solutions}
            Let $V$, $W$ and $X$ be real Hilbert spaces such that $W$ is a linear subspace of $X$, and $V$ is a linear subspace of $W$, i.e. $V \subset W \subset X$.
            \begin{itemize}
                \item[] X is the global space with inner product $(\cdot,\cdot)_X$ and the induced norm $||\cdot||_X$.
                \item[] W is the inertia space with inner product $(\cdot,\cdot)_W$ and the induced norm $||\cdot||_W$.
                \item[] V is the energy space with inner product $(\cdot,\cdot)_V$ and the induced norm $||\cdot||_V$.
            \end{itemize}
        \end{frame}

@@

                \item[] \textbf{A4} - The bilinear form $a$ is non-negative, symmetric and bounded on $V$, i.e. there exists a positive constant $K_a$ such that for $\displaystyle u,v \in V$, \[|a(u,v)| \leq K_a\Vert u \Vert_V \Vert v \Vert_V.\]
            \end{itemize}
        \end{frame}

@@

            This theorem allows for a solution of the abstract variational problem Problem GVar, if the assumptions \textbf{A1}-\textbf{A4} is satisfied and the initial values $u_0$ and $u_1$ are admissible. It is not always easy to verify that \eqref{bilinear_equation} is satisfied.
        \end{frame}

@@

                        \node at (6.95,1.25) {$b$};
                        \node at (6.78,0) {$h$};
                        \node at (3.2,-1.2) {$\ell = 1$};
                    \end{tikzpicture}
                    \captionof{figure}{Three-Dimensional Beam}
                \end{minipage}
                \hfill
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.6]
                        \draw[line width = 0.3mm] (0,1) -- (7,1);
                        \draw[line width = 0.3mm] (0,-1) -- (7,-1);
                        \draw[line width = 0.3mm] (7,-1) -- (7,1);
                        \draw[line width = 0.3mm] (0,-1) -- (0,1);

>>>>

\title{Finite element analysis of multi-dimensional and simplified models for beams}
\author{Rudi du Toit}
\date{\today}
\institute{University of Pretoria}

\newcommand{\footfullcitenonumber}[1]{
  \footnotetext{\fullcite{#1}}
}

@@

\section{Theory}
    \subsection{General vibration problem}
        \begin{frame}{General vibration problem}
            Let $V$, $W$ and $X$ be real Hilbert spaces such that $W$ is a linear subspace of $X$, and $V$ is a linear subspace of $W$, i.e. $V \subset W \subset X$.
            \begin{itemize}
                \item[] X is the global space with inner product $(\cdot,\cdot)_X$ and the induced norm $||\cdot||_X$.
                \item[] W is the inertia space with inner product $(\cdot,\cdot)_W$ and the induced norm $||\cdot||_W$.
                \item[] V is the energy space with inner product $(\cdot,\cdot)_V$ and the induced norm $||\cdot||_V$.
            \end{itemize}
            \only<2->{
            The bilinear forms $a$, $b$ and $c$ are defined as $a = (\cdot,\cdot)_X$,$b = (\cdot,\cdot)_V$ and $c = (\cdot,\cdot)_W$.
            }

            \only<3>{
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+a(u'(t),v)+b(u(t),v) = (f(t),v) \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.}

            \only<4->{
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+b(u(t),v) = (f(t),v) \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.}

            \footfullcitenonumber{VV02}
        \end{frame}

    \subsection{Existence and uniqueness of solutions}

        In [\cite{VV02}] the authors present results for the existence and uniqueness of solutions to the general vibration problem. The following assumptions are made for the existence results.

@@

                \item[] \textbf{A4} - The bilinear form $a$ is non-negative, symmetric and bounded on $V$, i.e. there exists a positive constant $K_a$ such that for $\displaystyle u,v \in V$, \[|a(u,v)| \leq K_a\Vert u \Vert_V \Vert v \Vert_V.\]
            \end{itemize}
            \footfullcitenonumber{VV02}
        \end{frame}

@@

            This theorem allows for a solution of the abstract variational problem Problem GVar, if the assumptions \textbf{A1}-\textbf{A4} is satisfied and the initial values $u_0$ and $u_1$ are admissible. It is not always easy to verify that \eqref{bilinear_equation} is satisfied.

            \footfullcitenonumber{VV02}
        \end{frame}

@@

                        \node at (6.95,1.25) {$b$};
                        \node at (6.78,0) {$h$};
                        \node at (3.2,-1.2) {$\ell = 1$};
                    \end{tikzpicture}
                    \captionof{figure}{Three-Dimensional Beam}
                \end{minipage}
                \hfill
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.6]
                        \draw[line width = 0.3mm] (0,1) -- (7,1);~
                        \draw[line width = 0.3mm] (0,-1) -- (7,-1);
                        \draw[line width = 0.3mm] (7,-1) -- (7,1);
                        \draw[line width = 0.3mm] (0,-1) -- (0,1);
